Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-08-20 mu/article/article.tex
Sending changes to BMS

\begin{abstract}
We present a theoretical study of spin-velocity injection (SVI) of a pure spin
current (PSC) induced by  a linearly polarized light that impinges normally on
the surface of  two 50\% hydrogenated noncentrosymmetric two-dimensional 2D
graphene structures. The first structure, hydrogenated at only one side,
labeled Up, also known as graphone, and the second, one labelled Alt, is
25\% hydrogenated at both sides. The hydrogenation opens an energy gap in both
structures. We analyze two possibilities: the first one is by fixing the spin
along a chosen direction, and the resulting SVI is calculated.
Secondly, we choose the SVI direction along
the surface plane, and calculate the resulting
spin orientation. This is done by changing the energy
$\hbar\omega$ and polarization angle $\alpha$  of the incoming  light. The
results are calculated within a full electronic band structure scheme using the
Density Functional Theory (DFT) in the Local Density Approximation (LDA). The
maxima of the spin-velocities are reached when $\hbar\omega=0.084$\,eV and
$\alpha=35^\circ$ for the Up structure, and $\hbar\omega=0.720$\,eV and
$\alpha=150^\circ$ for the Alt geometry. We find a speed of 668\,Km/s and
645\,Km/s for the Up and the Alt structures, respectively, when the spin points
perpendicularly to the surface. Also, the response is maximized by
fixing the spin-velocity direction 
along a high symmetry  axis, obtaining a speed of 688Km/s
with the spin pointing at $13^\circ$ from the surface normal, for the Up, 
and 906 Km/s and the spin pointing at $60^\circ$ from the surface normal, for
the Alt. These speed values are larger than those of bulk semiconductors,
such as CdSe and GaAs, thus making the hydrogenated graphene structures
excellent candidates for spintronics applications. 
\end{abstract}

@@

Graphene, an allotrope of carbon with hexagonal 2D lattice structure,
demonstrates properties such as fractional quantum Hall effect at room
temperature, excellent thermal transport properties, excellent
conductivity\cite{heerscheNat07} and strength \cite{geimNM07, reinaNL08,
novoselov2S07, balandinNL08}, being a perfect platform  for two-dimension (2D)
electronic systems; however, most electronic applications are disabled by the
absence of a semiconducting gap. Recent studies demonstrate that a narrow band
gap  can be opened in graphene by applying an electric field,\cite{zhangN09}
reducing the surface area,\cite{hanPRL07} or applying uniaxial
strain.\cite{niACSN08} Another possibility to open the gap is by doping; this
has been successfully achieved using nitrogen,\cite{weiNL2009} 
boron-nitrogen,\cite{guoIJ11} silicon,\cite{colettiPRB10} 
noble-metals,\cite{varykhalovPRB10} and hydrogen.\cite{eliasS09, guisingerNL09,
samarakoonACSN10} Depending on the percentage of hydrogenation and spatial
arrangements of the hydrogen-carbon bonds, hydrogenated graphene demonstrates
different structural configurations and a tunable electron gap, as it has been
proven in Ref. \onlinecite{shkrebtiiPSSC12}. 
In this paper, we present two 50\% hydrogenated graphene
noncentrosymmetric structures, both presenting a discernible band gap. The
first one, labelled as the Up structure, also known as
graphone,\cite{gmitraPRL13}  has
hydrogen atoms bonded to the carbon layer only on the upper side of the
structure; we consider here the magnetic isomer of graphone, with the so-called
``chair'' structure  shown in Fig. \ref{fig:up-struc}.
In contrast,  the Alt structure, shown in Fig. 
\ref{fig:alt-struc}, has hydrogen alternating on the upper and bottom sides of 
the carbon sheet.\cite{zapataPSB2016}

Both the Up and the Alt  structures
are noncentrosymmetric, and therefore,
they are good candidates in which SVI can be
induced. In this article, 
we address theoretically the SVI 
by one-photon absorption of linearly polarized light, analyzing
in our structures,
two possible scenarios of practical interest. 
The first case is by fixing the spin of the electrons along $z$, i.e. 
perpendicular 
to the surface plane, with the resulting velocity
directed along the surface of the
structures in the $xy$ plane. 
In the second case we fix the SVI velocity along the
$x$ or $y$ direction, and then, the resulting spin is directed outward of the
$xy$ plane.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=5cm]{figures/fig1}
    \caption{
Top (top panel) and side (bottom panel) views of the Up
      structure along with the 
      Cartesian $x$, $y$, and $z$ directions. The dark (light) spheres
      are the C (H) atoms. 
The primitive hexagonal unit cell is also shown.
{\color{red}Anatoli: we are confused about the primitive hexagonal
  unit cell shown. For instance, according to us, if we take C$_2$ and
apply the red arrow, it will end up in the site of C$_4$, but these C
atoms according to Table \ref{tab:up-unitcell} are not equivalent, or are they?}} 
    \label{fig:up-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=5cm]{figures/fig2}
    \caption{
Top (top panel) and side (bottom panel) views of the Alt
      structure along with the 
      Cartesian $x$, $y$, and$ z$ directions. The dark (light) spheres
      are the C (H) atoms. 
The primitive hexagonal unit cell is also shown.
}
    \label{fig:alt-struc}
\end{figure}

This paper is organized as follows. In Section \ref{sec:theory} we present the
formalism and the main expressions  that describe PSC and SVI. In Section
\ref{sec:results} we describe the numerical details  and  discuss the
corresponding SVI spectra for the Up and Alt structures. Finally, we summarize
our findings in Section \ref{sec:conclusions}.

@@

The operator that describes the electronic SVI is written as

@@

in this work.
To include the coherence, we follow Ref.~\onlinecite{nastosPRB07} and
use a multiple
scale approach that solves the equation of motion for the single
particle density matrix ${\rho}_{mn}({\mathbf k};t)$, leading to

@@

is the pseudotensor that describes the rate of change of the  PSC process in

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig3}
\caption{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ 
  {\it vs.} $\hbar\omega$, for the angles $\alpha$ that maximize the signal. The
  largest velocity are at the low energy regions of the spectra for the Alt and
  Up structures, becoming different from zero at the energy gap of each
  structure. In the high energy regions, the values of
  $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ are also very large compared to
  the 3D case of CdSe and GaAs, shown at the bottom panel.}
\label{fig:vab-str-comp}
\end{figure}

In Fig. \ref{fig:vab-str-comp}, we show $\mathcal{V}^{\mathrm{ab}}
(\omega,\alpha)$ {\it vs.} $\hbar\omega$ for the 
velocity and spin
directions
$\hat{\mathbf{a}}$ and $\hat{\mathbf{b}}$, and for the angle $\alpha$, at which
the signal is maximized, for the Up  and Alt structures, and for CdSe and GaAs
bulk systems, shown for comparison. As expected from
Eq.~\eqref{eq:mu}, 
${\cal V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ becomes different 
from zero right at the
corresponding energy gap of each system. For the 2D structures considered, the
spectrum contains two narrow energy regions with strong response, while for the
bulk systems, the spectra covers a rather wide energy range, but with a much
weaker response. For the Up structure, at $\mathrm{a}\mathrm{b}=yz$ and
$\alpha=35^\circ$ the response is maximized, which means that an incoming light
with its electric field polarized at 35$^\circ$ from the $x$ direction will
induce electrons to move along $y$, parallel to the surface, with their spin
polarized along $z$, perpendicular to the surface. The corresponding speeds
are the following: Right at the energy onset, ${\cal
V}^{yz}(\omega,\alpha)=668$\,Km/s, remains almost constant for 65 meV, and then
decreases to zero. A second region with high velocity is above 1.946 eV with
two, opposite in sign, maxima of the speed: ${\cal
V}^{yz}(\omega,\alpha)=266.3$ Km/s at $\hbar\omega=1.954$\,eV, and ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)=-241.4$ Km/s at
$\hbar\omega=1.958$\,eV; a positive (negative) ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ means that the electrons move
parallel (antiparallel) to the electric field. Likewise, for the Alt structure,
we also find that $\mathrm{a}\mathrm{b}=yz$ and $\alpha=150^\circ$ maximizes
the response, where two extreme values of ${\cal V}^{\mathrm{y}\mathrm{z}}$ are
found, one at  $\hbar\omega=0.720$\,eV of ${\cal
V}^{\mathrm{y}\mathrm{z}}=-711.9$ Km/s, and the other at
$\hbar\omega=0.911$\,eV of ${\cal V}^{\mathrm{y}\mathrm{z}}=-330.6$ Km/s.

For the bulk structures, we calculate ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$
from Eq.~\eqref{eq:vab-w} by simply using $\boldsymbol{\mu}_{\mathrm{max}}$.
For CdSe, we find that for $\hbar\omega=0.844$\,eV,
$\boldsymbol{\mu}_{\mathrm{max}}\to \mu^{zzzz}$, and ${\cal
V}^{\mathrm{z}\mathrm{z}}(\omega)=-59.0$ Km/s, and for GaAs at
$\hbar\omega=2.324$\,eV,
$\boldsymbol{\mu}_{\mathrm{max}}\to\mu^{\mathrm{aaaa}}$ and ${\cal
V}^{\mathrm{a}\mathrm{a}}(\omega,\alpha)=-28.7$\,Km/s, with $\mathrm{a}=x,y,z$.
For these bulk semiconductors, the $x$, $y$, and $z$ axis are taken along the
standard cubic unit cell directions, [100], [010], and [001], respectively. In
Table \ref{tab:vab-str-comp}, we compare ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ for the 2D structures considered and
bulk crystals. We stress that, as shown in the figure, for the 2D structures
have maxima in ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega;\alpha)$ are higher than
for the bulk crystals. In particular, the Alt structure demonstrates a ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega;\alpha)$ more than ten times larger than that
of CdSe and GaAs.
\begin{table}%[b]
\begin{tabular}{cccccc}
\hline
\multirow{2}{*}{Structure \quad} & 
System \quad & 
Pol. &
Energy & 
\multicolumn{2}{c}{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$}\\
\cline{5-6}
& type & Ang. & [eV] & $\mathrm{ab}$ \quad & [Km/s]\\
\hline
Up  & 2D   & 35    & 0.084  & $\mathrm{yz}$ &  660.5 \\
      &      &       & 1.954  & $\mathrm{yz}$ &  266.3 \\
      &      &       & 1.958  & $\mathrm{yz}$ & -241.4 \\
Alt & 2D   & 150   & 0.720  & $\mathrm{yz}$ & -711.9 \\
      &      &       & 0.911  & $\mathrm{yz}$ & -330.6 \\
CdSe  & bulk & --    & 0.844  & $\mathrm{zz}$ &  -59.0 \\
GaAs  & bulk & --    & 2.324  & $\mathrm{xx}$ &  -28.7 \\
\hline
\end{tabular}
\caption{Comparison of the reported maximum values of
$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ for the different structures and
their corresponding polarization angle $\alpha$ and energy $\hbar\omega$ . }
\label{tab:vab-str-comp}
\end{table}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig4}
\caption{Color online. For the Up structure, the top panel shows ${\cal
V}_{\sigma^\mathrm{z}}(\omega,\alpha)$ {\it vs.} $\hbar\omega$ and $\alpha$, and
the bottom panel shows $\gamma_{\sigma^\mathrm{z}}(\omega,\alpha)$ (right
scale, red line), and ${\cal
V}_{\sigma^\mathrm{z}}(\omega,\alpha)$ (left scale, black solid
line), {\it vs.} $\alpha$, for $\hbar\omega=0.084$\,eV, i.e. along
the ridge shown in the 3D plot. }
\label{fig:up-vsz-w1}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig5}
\caption{For the Up structure, the top panel shows ${\cal
V}_{\sigma^\mathrm{z}} (\omega,\alpha)$ {\it vs.} $\hbar\omega$ and $\alpha$, and the
bottom panel shows $\gamma_{\sigma^\mathrm{z}} (\omega,\alpha)$ (right scale,
red line, long dashes), and ${\cal V}_{\sigma^\mathrm{z}} (\omega,\alpha)$
(left scale, black solid line), {\it vs.} $\alpha$, for $\hbar\omega=1.954$\,eV, i.e.
along the highest ridge shown in the 3D plot. }
\label{fig:up-vsz-w2}
\end{figure}

@@

\begin{figure}[tb]
\centering
\includegraphics[width=\tama]{figures/fig6}
\caption{Color online. For the Alt structure, the top panel shows ${\cal
V}_{\sigma^\mathrm{z}} (\omega,\alpha)$ {\it vs.} $\hbar\omega$ and $\alpha$, and the
bottom panel shows $\gamma_{\sigma^\mathrm{z}} (\omega,\alpha)$ (right scale,
red line), and ${\cal V}_{\sigma^\mathrm{z}} (\omega,\alpha)$ (left scale,
black line), {\it vs.} $\alpha$, for $\hbar\omega=0.720$\,eV, i.e. along the ridge
shown in the 3D plot.}
\label{fig:alt-vsz}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig7}
\caption{
For the Up structure we show the velocity $\mathcal{V}_{\mathrm{a}}
(\omega,\alpha)$ (left scale, black line), the polar angle $\theta_{\mathrm{a}}
(\omega,\alpha)$ (right scale, red line), and  the azimuthal angle
$\varphi_{\mathrm{a}} (\omega,\alpha)$, (right scale, blue line), {\it vs.}
$\hbar\omega$, for $\alpha=35^\circ$, and $\mathrm{a}=x$ and $\mathrm{a}=y$.}
\label{fig:up-vab-comp-rtp-1}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig8}
\caption{For 
the Up structure we show the spin velocity $\mathcal{V}_{\mathrm{a}}
(\omega,\alpha)$ (left scale, black line), the polar angle $\theta_{\mathrm{a}}
(\omega,\alpha)$ (right scale, red line), and the azimuthal angle
$\varphi_{\mathrm{a}} (\omega,\alpha)$, (right scale, blue line), {\it vs.}
$\hbar\omega$, for $\alpha=35^\circ$, and $\mathrm{a}=x$ or $\mathrm{a}=y$. 
}
\label{fig:up-vx-vy-w2}
\end{figure}
In Fig. \ref{fig:up-vx-vy-w2}, we plot $\mathcal{V}_{\mathrm{a}}
(\omega,\alpha)$ {\it vs.} $\hbar\omega$, in the range where there two local maxima
with opposite sign at $\hbar\omega=1.954$\,eV and $\hbar\omega=1.957$\,eV
occur. The first is the largest of the two, with $\mathcal{V}_{\mathrm{x}}
(\omega,\alpha)=61.2$\,Km/s, $\theta_{\mathrm{x}}
(\omega,\alpha)=48.3^{\circ}$, and $\varphi_{\mathrm{x}}
(\omega,\alpha)=54.3^{\circ}$, for the electron moving along $x$; and
$\mathcal{V}_{\mathrm{y}} (\omega,\alpha)=293.2$\,Km/s, $\theta_{\mathrm{y}}
(\omega,\alpha)=49.8^{\circ}$, and $\varphi_{\mathrm{y}}
(\omega,\alpha)=51.9^{\circ}$ for the electron moving along $y$. For the peak
at $\hbar\omega=1.957$\,eV, we obtain $\theta_{\mathrm{x}} (\omega,\alpha) =
129.8^{\circ}$ and $\varphi_{\mathrm{x}} (\omega,\alpha) = 231.7^{\circ}$, with
$\mathcal{V}^{\mathrm{x}} (\omega,\alpha) = 54.6$\,Km/s and
$\theta_{\mathrm{y}}(\omega,\alpha) =129.3$; and
$\varphi_{\mathrm{y}}(\omega,\alpha) = 230.7$, with $\mathcal{V}^{\mathrm{y}}
(\omega,\alpha) = 263.7$\,Km/s. We remark that these angles are almost constant
for all the energy values across the peak of these two local maxima, for which
the spin is directed upward in the first quadrant of the $xy$ plane when
the electron moves along either $x$ or
$y$ directions.

@@

\begin{figure}[tb]
\centering
\includegraphics[width=\tama]{figures/fig9}
\caption{For the Alt structure we show  the velocity $\mathcal{V}_{\mathrm{a}}
(\omega,\alpha)$ (left scale, black line), the polar angle $\theta_{\mathrm{a}}
(\omega,\alpha)$ (right scale, red line), and the azimuthal angle
$\varphi_{\mathrm{a}} (\omega,\alpha)$, (right scale, blue line), {\it vs.}
$\hbar\omega$, for $\alpha=150^\circ$, and $\mathrm{a}=x$ or $\mathrm{a}=y$.
}
\label{fig:alt-vx-vy-w1}
\end{figure}

>>>>

\begin{abstract}
We present a theoretical study of spin-velocity injection (SVI) of a pure spin
current (PSC) induced by  a linearly polarized light that impinges normally on
the surface of  two 50\% hydrogenated noncentrosymmetric two-dimensional 2D
graphene structures. The first structure, hydrogenated at only one side,
labeled Up, also known as graphone, and the second, labelled Alt, is
25\% hydrogenated at both sides. The hydrogenation opens an energy gap in both
structures. We analyze two possibilities: the first, the spin is fixed
along a chosen direction, and the resulting SVI is calculated.
In the second, we choose the SVI direction along
the surface plane, and calculate the resulting
spin orientation. This is done by changing the energy
$\hbar\omega$ and polarization angle $\alpha$  of the incoming  light. The
results are calculated within a full electronic band structure scheme using the
Density Functional Theory (DFT) in the Local Density Approximation (LDA). The
maxima of the spin-velocities are reached when $\hbar\omega=0.084$\,eV and
$\alpha=35^\circ$ for the Up structure, and $\hbar\omega=0.720$\,eV and
$\alpha=150^\circ$ for the Alt geometry. We find a speed of 668\,Km/s and
645\,Km/s for the Up and the Alt structures, respectively, when the spin points
perpendicularly to the surface. Also, the response is maximized by
fixing the spin-velocity direction 
along a high symmetry  axis, obtaining a speed of 688Km/s
with the spin pointing at $13^\circ$ from the surface normal, for the Up, 
and 906 Km/s and the spin pointing at $60^\circ$ from the surface normal, for
the Alt system. These speed values are  of order of magnitude larger than those of bulk semiconductors,
such as CdSe and GaAs, thus making the hydrogenated graphene structures
excellent candidates for spintronics applications. 
\end{abstract}

@@

Graphene, an allotrope of carbon with hexagonal 2D lattice structure,
demonstrates properties such as fractional quantum Hall effect at room
temperature, excellent thermal transport properties, excellent
conductivity\cite{heerscheNat07} and strength \cite{geimNM07, reinaNL08,
novoselov2S07, balandinNL08}, being a perfect platform  for two-dimensional (2D)
electronic systems; however, numerous important electronic applications are disabled by the
absence of a semiconducting gap. Recent studies demonstrate that a narrow band
gap  can be opened in graphene by applying an electric field,\cite{zhangN09}
reducing the surface area,\cite{hanPRL07} or applying uniaxial
strain.\cite{niACSN08} Another possibility to open the gap is by doping; this
has been successfully achieved using nitrogen,\cite{weiNL2009} 
boron-nitrogen,\cite{guoIJ11} silicon,\cite{colettiPRB10} 
noble-metals,\cite{varykhalovPRB10} and hydrogen.\cite{eliasS09, guisingerNL09,
samarakoonACSN10} Depending on the percentage of hydrogenation and spatial
arrangements of the hydrogen-carbon bonds, hydrogenated graphene demonstrates
different structural configurations and a tunable electron gap, as it has been
proven in Ref. \onlinecite{shkrebtiiPSSC12}. 
In this paper, we offer two 50\% hydrogenated graphene
noncentrosymmetric structures, both demonstrating a discernible band gap. The
first one, labelled as the Up structure, also known as
graphone,\cite{gmitraPRL13}  has
hydrogen atoms bonded to the carbon layer only on the upper side of the
structure; we consider here the magnetic isomer of graphone, with the so-called
``chair'' structure  shown in Fig. \ref{fig:up-struc}.
In contrast,  the Alt structure, shown in Fig. 
\ref{fig:alt-struc}, has hydrogen alternating on the upper and bottom sides of 
the carbon sheet.\cite{zapataPSB2016}

Both the Up and the Alt  structures
are noncentrosymmetric, and therefore,
they are good candidates in which SVI can be
induced. In this article, 
we address theoretically the spin-velocity injection
by one-photon absorption of linearly polarized light, analyzing
in our structures,
two possible scenarios of practical interest. 
The first case is by fixing the spin of the electrons along $z$, i.e. 
perpendicular 
to the surface plane, with the resulting velocity
directed along the surface of the
structures in the $xy$ plane. 
In the second case we fix the SVI velocity along the
$x$ or $y$ direction, and then, the resulting spin is directed outward of the
$xy$ plane.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=5cm]{figures/fig1}
    \caption{(color online). Top (top panel) and side (bottom panel) views of
the Up structure along with the Cartesian $x$, $y$, and $z$ directions. The
dark (light) spheres are the C (H) atoms. The primitive hexagonal unit cell is
also shown.}
    \label{fig:up-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=5cm]{figures/fig2}
    \caption{(color online). Top (top panel) and side (bottom panel) views of
the Alt structure along with the Cartesian $x$, $y$, and$ z$ directions. The
dark (light) spheres are the C (H) atoms. The primitive rectangular unit cell
is also shown. }
    \label{fig:alt-struc}
\end{figure}

This paper is organized as follows. In Section \ref{sec:theory} we outline the
formalism and the main expressions  that describe PSC and SVI. In Section
\ref{sec:results} we describe the numerical details  and  discuss the
corresponding SVI spectra for the Up and Alt structures. Finally, we summarize
our findings in Section \ref{sec:conclusions}.

@@

To calculate the velocity of the spin injection
${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$ along the direction $\hat{\mathbf{a}}$, at
which the spin moves in a
polarized state along direction $\hat{\mathbf{b}}$,
we start with the operator that describes the electronic SVI, written as

@@

in this work. To include the coherence, we follow Ref.~\onlinecite{nastosPRB07}
and use a multiple scale approach that solves the equation of motion for the
single particle density matrix ${\rho}_{mn}({\mathbf k};t)$, leading to
{\color{red} just for comments to think about, not to introduce the suggestions
right away: above we have $\rho_{mn}$, below $\rho_{cc’}$ time variable has
disappeared below. And $\rho_{cc’}$ contains a and b labels, which are not
Cartesian}

@@

is the pseudotensor that describes the rate of change of the PSC in

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig3}
\caption{(color online).  Spin velocity injection
  $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ {\it vs.} photon energy
  $\hbar\omega$, for the angles $\alpha$ that maximize the signal. The largest
  velocity are at the low energy regions of the spectra for the Alt and Up
  structures, becoming different from zero at the energy gap of each structure.
  In the high energy regions, the values of
  $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ are also very large compared to
  the 3D case of CdSe and GaAs, shown at the bottom panel.}
\label{fig:vab-str-comp}
\end{figure}

In Fig. \ref{fig:vab-str-comp}, we show $\mathcal{V}^{\mathrm{ab}}
(\omega,\alpha)$ {\it vs.} $\hbar\omega$ for the 
velocity and spin
directions
$\hat{\mathbf{a}}$ and $\hat{\mathbf{b}}$, and for the angle $\alpha$, at which
the signal is maximized, for the Up  and Alt structures, and for CdSe and GaAs
bulk systems, shown for comparison. As expected from
Eq.~\eqref{eq:mu}, 
${\cal V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ starts rising 
from zero right at the
corresponding energy gap of each system. For the 2D structures considered, the
spectrum contains two narrow energy regions with strong response, while for the
bulk systems, the spectra covers a rather wide energy range, but with a much
weaker response. For the Up structure, at $\mathrm{a}\mathrm{b}=yz$ and
$\alpha=35^\circ$ the response is maximized, which means that an incoming light
with its electric field polarized at 35$^\circ$ from the $x$ direction will
induce electrons to move along $y$, parallel to the surface, with their spin
polarized along $z$, perpendicular to the surface. Right at the energy onset, ${\cal
V}^{yz}(\omega,\alpha)=668$\,Km/s, remains almost constant for 65 meV, and then
decreases to zero. A second region with high velocity is above 1.946 eV with
two, opposite in sign, maxima of the speed: ${\cal
V}^{yz}(\omega,\alpha)=266.3$ Km/s at $\hbar\omega=1.954$\,eV, and ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)=-241.4$ Km/s at
$\hbar\omega=1.958$\,eV; a positive (negative) ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ means that the electrons move
parallel (antiparallel) to the electric field. Likewise, for the Alt structure,
we also find that $\mathrm{a}\mathrm{b}=yz$ and $\alpha=150^\circ$ maximizes
the response, where two extreme values of ${\cal V}^{\mathrm{y}\mathrm{z}}$ are
found, one at  $\hbar\omega=0.720$\,eV of ${\cal
V}^{\mathrm{y}\mathrm{z}}=-711.9$ Km/s, and the other at
$\hbar\omega=0.911$\,eV of ${\cal V}^{\mathrm{y}\mathrm{z}}=-330.6$ Km/s.

For the bulk structures, we calculate ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$
from Eq.~\eqref{eq:vab-w} by simply using $\boldsymbol{\mu}_{\mathrm{max}}$.
For CdSe, we find that for $\hbar\omega=0.844$\,eV,
$\boldsymbol{\mu}_{\mathrm{max}}\to \mu^{zzzz}$, and ${\cal
V}^{\mathrm{z}\mathrm{z}}(\omega)=-59.0$ Km/s, and for GaAs at
$\hbar\omega=2.324$\,eV,
$\boldsymbol{\mu}_{\mathrm{max}}\to\mu^{\mathrm{aaaa}}$ and ${\cal
V}^{\mathrm{a}\mathrm{a}}(\omega,\alpha)=-28.7$\,Km/s, with $\mathrm{a}=x,y,z$.
For these bulk semiconductors, the $x$, $y$, and $z$ axis are taken along the
standard cubic unit cell directions, [100], [010], and [001], respectively. In
Table \ref{tab:vab-str-comp}, we compare ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ for the 2D structures considered and
bulk crystals. We stress that, as shown in the figure, for the 2D structures
have maxima in ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega;\alpha)$ are higher than
for the bulk crystals more than order of magnitude. In particular, the Alt structure demonstrates a ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega;\alpha)$ about 12 times larger than that
of CdSe and GaAs.

{\color{red} Finished here; A.}
\begin{table}%[b]
\begin{tabular}{cccccc}
\hline
\multirow{2}{*}{Structure \quad} & 
System \quad & 
Pol. &
Energy & 
\multicolumn{2}{c}{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$}\\
\cline{5-6}
& type & Ang. & [eV] & $\mathrm{ab}$ \quad & [Km/s]\\
\hline
Up  & 2D   & 35    & 0.084  & $\mathrm{yz}$ &  660.5 \\
      &      &       & 1.954  & $\mathrm{yz}$ &  266.3 \\
      &      &       & 1.958  & $\mathrm{yz}$ & -241.4 \\
Alt & 2D   & 150   & 0.720  & $\mathrm{yz}$ & -711.9 \\
      &      &       & 0.911  & $\mathrm{yz}$ & -330.6 \\
CdSe  & bulk & --    & 0.844  & $\mathrm{zz}$ &  -59.0 \\
GaAs  & bulk & --    & 2.324  & $\mathrm{xx}$ &  -28.7 \\
\hline
\end{tabular}
\caption{Comparison of the reported maximum values of
$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ for the different structures and
their corresponding polarization angle $\alpha$ and energy $\hbar\omega$ . }
\label{tab:vab-str-comp}
\end{table}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig4}
\caption{(color online). Color online. For the Up structure, the top panel
shows ${\cal V}_{\sigma^\mathrm{z}}(\omega,\alpha)$ {\it vs.} $\hbar\omega$ and
$\alpha$, and the bottom panel shows
$\gamma_{\sigma^\mathrm{z}}(\omega,\alpha)$ (right scale, red line), and ${\cal
V}_{\sigma^\mathrm{z}}(\omega,\alpha)$ (left scale, black solid line), {\it
vs.} $\alpha$, for $\hbar\omega=0.084$\,eV, i.e. along the ridge shown in the
3D plot. }
\label{fig:up-vsz-w1}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig5}
\caption{(color online). For the Up structure, the top panel shows ${\cal
V}_{\sigma^\mathrm{z}} (\omega,\alpha)$ {\it vs.} $\hbar\omega$ and $\alpha$,
and the bottom panel shows $\gamma_{\sigma^\mathrm{z}} (\omega,\alpha)$ (right
scale, red line, long dashes), and ${\cal V}_{\sigma^\mathrm{z}}
(\omega,\alpha)$ (left scale, black solid line), {\it vs.} $\alpha$, for
$\hbar\omega=1.954$\,eV, i.e. along the highest ridge shown in the 3D plot. }
\label{fig:up-vsz-w2}
\end{figure}

@@

\begin{figure}[tb]
\centering
\includegraphics[width=\tama]{figures/fig6}
\caption{(color online). Color online. For the Alt structure, the top panel
shows ${\cal V}_{\sigma^\mathrm{z}} (\omega,\alpha)$ {\it vs.} $\hbar\omega$
and $\alpha$, and the bottom panel shows $\gamma_{\sigma^\mathrm{z}}
(\omega,\alpha)$ (right scale, red line), and ${\cal V}_{\sigma^\mathrm{z}}
(\omega,\alpha)$ (left scale, black line), {\it vs.} $\alpha$, for
$\hbar\omega=0.720$\,eV, i.e. along the ridge shown in the 3D plot.}
\label{fig:alt-vsz}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig7}
\caption{(color online). For the Up structure we show the velocity
$\mathcal{V}_{\mathrm{a}} (\omega,\alpha)$ (left scale, black line), the polar
angle $\theta_{\mathrm{a}} (\omega,\alpha)$ (right scale, red line), and  the
azimuthal angle $\varphi_{\mathrm{a}} (\omega,\alpha)$, (right scale, blue
line), {\it vs.} $\hbar\omega$, for $\alpha=35^\circ$, and $\mathrm{a}=x$ and
$\mathrm{a}=y$.}
\label{fig:up-vab-comp-rtp-1}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig8}
\caption{(color online). For the Up structure we show the spin velocity
$\mathcal{V}_{\mathrm{a}} (\omega,\alpha)$ (left scale, black line), the polar
angle $\theta_{\mathrm{a}} (\omega,\alpha)$ (right scale, red line), and the
azimuthal angle $\varphi_{\mathrm{a}} (\omega,\alpha)$, (right scale, blue
line), {\it vs.} $\hbar\omega$, for $\alpha=35^\circ$, and $\mathrm{a}=x$ or
$\mathrm{a}=y$. }
\label{fig:up-vx-vy-w2}
\end{figure}
In Fig. \ref{fig:up-vx-vy-w2}, we plot $\mathcal{V}_{\mathrm{a}}
(\omega,\alpha)$ {\it vs.} $\hbar\omega$, in the range where there two local maxima
with opposite sign at $\hbar\omega=1.954$\,eV and $\hbar\omega=1.957$\,eV
occur. The first is the largest of the two, with $\mathcal{V}_{\mathrm{x}}
(\omega,\alpha)=61.2$\,Km/s, $\theta_{\mathrm{x}}
(\omega,\alpha)=48.3^{\circ}$, and $\varphi_{\mathrm{x}}
(\omega,\alpha)=54.3^{\circ}$, for the electron moving along $x$; and
$\mathcal{V}_{\mathrm{y}} (\omega,\alpha)=293.2$\,Km/s, $\theta_{\mathrm{y}}
(\omega,\alpha)=49.8^{\circ}$, and $\varphi_{\mathrm{y}}
(\omega,\alpha)=51.9^{\circ}$ for the electron moving along $y$. For the peak
at $\hbar\omega=1.957$\,eV, we obtain $\theta_{\mathrm{x}} (\omega,\alpha) =
129.8^{\circ}$ and $\varphi_{\mathrm{x}} (\omega,\alpha) = 231.7^{\circ}$, with
$\mathcal{V}^{\mathrm{x}} (\omega,\alpha) = 54.6$\,Km/s and
$\theta_{\mathrm{y}}(\omega,\alpha) =129.3$; and
$\varphi_{\mathrm{y}}(\omega,\alpha) = 230.7$, with $\mathcal{V}^{\mathrm{y}}
(\omega,\alpha) = 263.7$\,Km/s. We remark that these angles are almost constant
for all the energy values across the peak of these two local maxima, for which
the spin is directed upward in the first quadrant of the $xy$ plane when
the electron moves along either $x$ or
$y$ directions.

@@

\begin{figure}[tb]
\centering
\includegraphics[width=\tama]{figures/fig9}
\caption{(color online). For the Alt structure we show  the velocity
$\mathcal{V}_{\mathrm{a}} (\omega,\alpha)$ (left scale, black line), the polar
angle $\theta_{\mathrm{a}} (\omega,\alpha)$ (right scale, red line), and the
azimuthal angle $\varphi_{\mathrm{a}} (\omega,\alpha)$, (right scale, blue
line), {\it vs.} $\hbar\omega$, for $\alpha=150^\circ$, and $\mathrm{a}=x$ or
$\mathrm{a}=y$. }
\label{fig:alt-vx-vy-w1}
\end{figure}
